\documentclass[12pt,a4paper]{article}
\usepackage[latin1]{inputenc}
\usepackage{amsmath}
\usepackage{amsfonts}
\usepackage{amssymb}
\usepackage{graphicx}
\newcounter{pubcounter}
\begin{document}
	List of Publications:
	\begin{itemize}
			\item{Published and Accepted}
		\begin{enumerate}
			\item{ Li, S., Wang, Z., Yau, S. S. T., Zhang, Z.,Tensor train method for high-dimensional nonlinear filtering problems  \textsl{(IEEE TAC, to appear)}}
			\item{Wang, Z., Xin, J., Zhang, Z., DeepParticle: learning invariant measure by a deep neural network minimizing Wasserstein distance on data generated from an interacting particle method, Journal of Computational Physics (2022): 111309.}
			\item{Wang, Z., Xin, J., Zhang, Z., Computing effective diffusivities in 3D time-dependent chaotic flows with a convergent Lagrangian numerical method, ESAIM: M2AN 56 (2022) 1521?1544}
			\item{ Lyu, J., Wang, Z., Xin, J., Zhang, Z., A convergent interacting particle method and computation of KPP front speeds in chaotic flows, SIAM Journal on Numerical Analysis, 2022, 60(3): 1136-1167}
			\item{Wang, Z., Xin, J., Zhang, Z., Sharp uniform in time error estimate on a stochastic structure-preserving Lagrangian method and computation of effective diffusivity in 3D chaotic flows, Multiscale Model and Simulation, 19 (2021), no. 3, 1167?1189}
			\item{ Lyu, J., Wang, Z., Xin, J., Zhang, Z., Convergence of stochastic structure-preserving schemes for computing effective diffusivity in random flows, SIAM Journal on Numerical Analysis, 58 (2020), no. 5, 3040?3067. .}
			\item{Wang, Z., Zhang, Z., A new mesh-free method for PDE with discontinuous coeffcients using the deep learning approach,  Journal of Computational Physics (2020): 108963.}
			\item{Wang, Z. Luo, X., Yau, S. S. T., Zhang, Z., Proper orthogonal decomposition method to nonlinear filtering problems in medium-high dimension, IEEE Transactions on Automatic Control, 65 (2020), no. 4, 1613?1624.}
			\item{Wang, Z., Xin, J., Zhang, Z.,  Computing Effective Diffusivity of Chaotic and Stochastic Flows Using Structure-Preserving Schemes. SIAM Journal on Numerical Analysis, 56(4), 2322-2344.}
			\setcounter{pubcounter}{\value{enumi}}
		\end{enumerate}
		\item{Preprints and Under-preparation}
		\begin{enumerate}
			\setcounter{enumi}{\value{pubcounter}}
			\item{Wang, Z., Zhang W., Zhang Z., A data-driven model reduction method for parabolic inverse source problems and its convergence analysis \textsl{(arXiv:2110.07676)}}
			\item{Wang, Z., T. Hou, Zhang, Z.A class of robust numerical methods for solving dynamical systems with multiple time scales  \textsl{(arXiv:1909.04289)}}
			\item{Bal, G., Hoskins, J.G., Wang Z., Transport in Dirac models for topological insulators \textsl{(arXiv:2206.08847)}}
			\item{Wang, Z., Xin, J., Zhang Z.,A DeepParticle method for learning and generating aggregation patterns in multi-dimensional Keller-Segel chemotaxis systems  \textsl{(arXiv:2209.00109)}}
			\item{Cui, T., Wang, Z., Zhang, Z., A variational neural network approach for glacier modelling with nonlinear rheology \textsl{(arXiv:2209.02088)}}
			\item{Chen, B., Bal, G., Wang, Z. One Dimensional Mixed Type Generalized Kimura Type Operator}
			\item{Xie, Y., Wang, Z., Zhang, Z., Mini-batch interior point method for computing optimal transportation problems and its convergence analysis}
			\item{Wang, Z., Zhang W., Zhang Z., Stochastic convergence of regularized solutions for backward heat conduction problems}
			\item{Non-convergence tests for trajectory averages of ergodic Markov chains and diffusions}
			\setcounter{pubcounter}{\value{enumi}}
		\end{enumerate}
		\item{Dissertations}
		\begin{enumerate}
			\setcounter{enumi}{\value{pubcounter}}
			\item Robust Lagrangian Numerical Schemes in Computing Effective Diffusivities for Chaotic and Random Flows, \textsl{Ph.D. Thesis, advisor: Prof. Zhiwen Zhang at HKU}
			\item Convergence analysis of strong approximation to stochastic differential equation, \textsl{Bachelor Thesis, advisor: Prof. Espen Robstad Jakobsen at ENS}
			\setcounter{pubcounter}{\value{enumi}}
		\end{enumerate}
	\end{itemize}
\end{document}